\section{Step 6: IoT Level Specification}
The Smart Mosque Management System is structured across five IoT levels, following a standard layered architecture. Each layer has a defined responsibility and communicates with adjacent layers.

\subsection{Level 1 -- Perception Layer}
\textbf{Role:} Direct interaction with the physical environment.

\textbf{Sensors}
\begin{table}[htbp]
\centering
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}L{4.1cm} L{4.1cm} L{5.5cm}}
\toprule
\textbf{Device} & \textbf{Placement} & \textbf{Measurement} \\
\midrule
PIR Motion Sensor & Main entrance, exit, hall zone & Entry/exit events, zone presence \\
Water Flow Sensor & Wudu tap inlet pipe & Flow rate (L/min), pulse output \\
PIR Motion Sensor (Wudu) & Ablution area & Presence near taps \\
IAQ Sensor & Prayer hall near ventilation & CO$_2$, temperature, humidity \\
\bottomrule
\end{tabular*}
\caption{Perception layer sensors}
\end{table}

\textbf{Actuators}
\begin{table}[htbp]
\centering
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}L{4.1cm} L{4.1cm} L{5.5cm}}
\toprule
\textbf{Device} & \textbf{Placement} & \textbf{Controlled Action} \\
\midrule
Smart Relay (Lights) & Per hall zone & ON/OFF lighting \\
Smart Relay (Fans/AC) & Per hall zone, wudu area & ON/OFF ventilation \\
Solenoid Valve & Wudu tap supply line & Open/close water flow \\
\bottomrule
\end{tabular*}
\caption{Perception layer actuators}
\end{table}

\subsection{Level 2 -- Network Layer}
\textbf{Role:} Local connectivity and data transport.
\begin{itemize}
  \item Hardware: Local Wi-Fi access point for ESP32 nodes and local server
  \item Protocol: IEEE 802.11 Wi-Fi with WPA2-PSK
  \item Topology: Star
  \item Segmentation: IoT network isolated from guest/user devices
\end{itemize}

Key decision: Core control loops are local and do not depend on internet connectivity.

\subsection{Level 3 -- Edge Processing Layer}
\textbf{Role:} Real-time control and decision-making.

\textbf{Hardware:} ESP32 microcontrollers per functional zone.

\begin{table}[htbp]
\centering
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}L{4.6cm} L{8.8cm}}
\toprule
\textbf{Function} & \textbf{Description} \\
\midrule
Sensor polling & Reads PIR, IAQ, and flow sensor inputs at configured intervals \\
Occupancy counting & Increments/decrements people count based on entry/exit events \\
Prayer window evaluation & Compares current time against cached prayer schedule \\
Energy control logic & Combines occupancy and prayer window state for relays \\
Water control logic & Detects anomalies and closes valve when required \\
Air quality control & Polls IAQ every 60s and activates fans on threshold breach \\
Alert generation & Publishes threshold breach events to data layer \\
\bottomrule
\end{tabular*}
\caption{Edge processing responsibilities}
\end{table}

Failsafe: If server connectivity is lost, edge controllers continue autonomous operation.

\subsection{Level 4 -- Data Accumulation Layer}
\textbf{Role:} Aggregate processed data, store logs, and serve dashboard queries.
\begin{itemize}
  \item CPT simulation backend: Registration Server
  \item Real deployment option: local server (e.g., lightweight REST + database)
  \item Retention: rolling 30-day window with optional archival
\end{itemize}

\subsection{Level 5 -- Application Layer}
\textbf{Role:} Human-facing dashboard for monitoring and control.
\begin{itemize}
  \item Web dashboard on local network
  \item Single admin role
  \item Functions: live monitoring, alert display, schedule view, manual override
\end{itemize}

\subsection{Level Interaction Summary}
\begin{verbatim}
[Sensors / Actuators] (Level 1)
      | GPIO / I2C / UART
      v
[ESP32 Edge Controller] (Level 2+3) <--- [Prayer Time API]
      | Wi-Fi (WPA2-PSK)
      v
[Local Server / Registration Server] (Level 4)
      | Local LAN
      v
[Admin Web Dashboard] (Level 5)
\end{verbatim}
